### Load libraries and files as tibles
```{r load libraries and data, include=FALSE}
library(tidyverse)
library(patchwork)
library(knitr)

directory <- "/Users/anaso/OneDrive - BABRAHAM/experiments/OOPS_AB031_CN/day_2/tiff_for_cell_profiler"
csv_files <- list.files(path = directory, pattern = ".csv", full.names = TRUE)
data_list <- map(csv_files, read_csv)

for (i in seq_along(csv_files)) {
  # Extract the file name without extension
  file_name <- tools::file_path_sans_ext(basename(csv_files[i]))
  # Assign the tibble to a variable with the file name
  assign(file_name, data_list[[i]])
}
```


### Remove prefix from file titles
```{r read_the_files, echo = T, message=FALSE}

# (optional) rename variables by removing the experiment prefix
var_names <- ls(pattern = "OOPS_AB031_")
for (old_name in var_names) {
  # Generate the new name by removing "prefix"
  new_name <- gsub("OOPS_AB031_", "", old_name)
  
  # Use assign to rename the variable
  assign(new_name, get(old_name))
  rm(list = old_name)
}
```

### Create metadata
```{r infer metadata from filenames, include=FALSE}

Metadata <- Image[, c("FileName_G3Bp1", "ImageNumber")]

Metadata <- Metadata |>
  mutate(Stressor = case_when(
    str_detect(FileName_G3Bp1, "ars") ~ "NaAsO2 treated",
    str_detect(FileName_G3Bp1, "untr") ~ "Untreated",
    TRUE ~ "ERROR"
  ))

Metadata <- Metadata |> 
  mutate(CellLine = case_when(
    str_detect(FileName_G3Bp1, "WT") ~ "WT v-abl",
    str_detect(FileName_G3Bp1, "8A6") ~ "Clone 8A6",
    str_detect(FileName_G3Bp1, "8B10") ~ "Clone 8B10",
    str_detect(FileName_G3Bp1, "8D6") ~ "Clone 8D6",
    str_detect(FileName_G3Bp1, "8D9") ~ "Clone 8D9",
    str_detect(FileName_G3Bp1, "8E10") ~ "Clone 8E10",
    str_detect(FileName_G3Bp1, "8F5") ~ "Clone 8F5",
    str_detect(FileName_G3Bp1, "8G5") ~ "Clone 8G5",
    str_detect(FileName_G3Bp1, "8H6") ~ "Clone 8H6",
    TRUE ~ "ERROR"
  ))


Metadata <- Metadata[c(2,3,4)]
```


### Filter data within Image and calculate number of SGs per cell
```{r filter total SGs per total cells, echo=FALSE}

# choose only relevant columns
Image_filtered <- Image[, c("ImageNumber", "Count_SG_G3Bp1_in_cells" , "Count_SG_G3Bp2_in_cells" , "Count_G3Bp1_G3Bp2_SG" , "Count_cell")]

# join with metadata
Image_filtered <- Image_filtered %>% left_join(Metadata, by = "ImageNumber")
Image_filtered <- Image_filtered |> mutate(SG_per_cell_G1 = Count_SG_G3Bp1_in_cells/Count_cell)
Image_filtered <- Image_filtered |> mutate(SG_per_cell_G2 = Count_SG_G3Bp2_in_cells/Count_cell)
Image_filtered <- Image_filtered |> mutate(SG_per_cell_G1G2 = Count_G3Bp1_G3Bp2_SG/Count_cell)
```

### Summarise data per cell_line and Stressor
```{r summarise data per cell_line and stressor based on number of SGs per cell}

Image_filtered_summary <- Image_filtered %>%
  group_by(CellLine, Stressor) %>%
  summarise(
    n = n(),                     # Count observations in each group
    mean_SG_per_cell_G1 = mean(SG_per_cell_G1),    # Calculate mean Ratio
    sd_Ratio_G3Bp1 = sd(SG_per_cell_G1),        # Calculate standard deviation
    se_Ratio_G3Bp1 = sd(SG_per_cell_G1) / sqrt(n()), # Calculate standard error
    mean_Ratio_G3Bp2 = mean(SG_per_cell_G2),    # Calculate mean Ratio
    sd_Ratio_G3Bp2 = sd(SG_per_cell_G2),        # Calculate standard deviation
    se_Ratio_G3Bp2 = sd(SG_per_cell_G2) / sqrt(n()), # Calculate standard error
    mean_Ratio_G3Bp1_G3Bp2 = mean(SG_per_cell_G1G2),    # Calculate mean Ratio
    sd_Ratio_G3Bp1_G3Bp2 = sd(SG_per_cell_G1G2),        # Calculate standard deviation
    se_Ratio_G3Bp1_G3Bp2 = sd(SG_per_cell_G1G2) / sqrt(n()) # Calculate standard error
  ) %>%
  ungroup()
```

### Plots of average number of SGs per total number of cells
```{r total SGs per total cells, echo=FALSE}

#change order of labels
Image_filtered_summary$Stressor <- factor(Image_filtered_summary$Stressor, 
                                  levels = c("Untreated", "NaAsO2 treated"))
# set colours
manual_colors <- c("NaAsO2 treated" = "#00bfc4", "Untreated" = "#F8766D")  

#Number of SGs per cell plot - G3BP1
average_SG_G3Bp1_plot <- Image_filtered_summary |>
  ggplot(aes(x = CellLine, y = mean_SG_per_cell_G1, fill = Stressor)) + 
  geom_bar(stat = "identity", position = position_dodge(width = 0.9)) +  
  geom_errorbar(
    aes(ymin = mean_SG_per_cell_G1 - se_Ratio_G3Bp1, ymax = mean_SG_per_cell_G1 + se_Ratio_G3Bp1),
    width = 0.2, # Width of the error bar caps
    position = position_dodge(width = 0.9) # Ensure error bars align with dodged bars
  ) +
  labs(
    title = "Average number of SGs per cell (based on G3Bp1 signal)",
    y = "Average number of SGs per cell",
  ) + theme_minimal() + 
   scale_fill_manual(values = manual_colors) + scale_y_continuous(limits = c(0, 1.8)) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1), legend.position = "bottom", legend.title = element_blank(), axis.title.x = element_blank()) + theme(
         panel.background = element_rect(fill='transparent'),
         plot.background = element_rect(fill='transparent', color=NA),
         panel.grid.major = element_blank(),
         panel.grid.minor = element_blank(),
         legend.background = element_rect(fill='transparent'),
         legend.box.background = element_rect(fill='transparent')
       )

#Number of SGs per cell plot - G3Bp2
average_SG_G3Bp2_plot <- Image_filtered_summary |>
  ggplot(aes(x = CellLine, y = mean_Ratio_G3Bp2, fill = Stressor)) + 
  geom_bar(stat = "identity", position = position_dodge(width = 0.9)) +  
  geom_errorbar(
    aes(ymin = mean_Ratio_G3Bp2 - se_Ratio_G3Bp2, ymax = mean_Ratio_G3Bp2 + se_Ratio_G3Bp2),
    width = 0.2, # Width of the error bar caps
    position = position_dodge(width = 0.9) # Ensure error bars align with dodged bars
  ) +
  labs(
    title = "Average number of SGs per cell (based on G3Bp2 signal)",
    y = "Average number of SGs per cell",
  ) + theme_minimal() + 
   scale_fill_manual(values = manual_colors) + scale_y_continuous(limits = c(0, 4)) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1), legend.position = "bottom", legend.title = element_blank(), axis.title.x = element_blank()) + theme(
         panel.background = element_rect(fill='transparent'),
         plot.background = element_rect(fill='transparent', color=NA),
         panel.grid.major = element_blank(),
         panel.grid.minor = element_blank(),
         legend.background = element_rect(fill='transparent'),
         legend.box.background = element_rect(fill='transparent')
       )

#Number of SGs per cell plot - G3Bp1/G3Bp2
average_SG_G3Bp2_G3Bp1_plot <- Image_filtered_summary  |>
  ggplot(aes(x = CellLine, y = mean_Ratio_G3Bp1_G3Bp2, fill = Stressor)) + 
  geom_bar(stat = "identity", position = position_dodge(width = 0.9)) +  
  geom_errorbar(
    aes(ymin = mean_Ratio_G3Bp1_G3Bp2 - se_Ratio_G3Bp1_G3Bp2, ymax = mean_Ratio_G3Bp1_G3Bp2 + se_Ratio_G3Bp1_G3Bp2),
    width = 0.2, # Width of the error bar caps
    position = position_dodge(width = 0.9) # Ensure error bars align with dodged bars
  ) +
  labs(
    title = "Average number of SGs per cell (based on G3Bp1 and G3Bp2 signal)",
    y = "Average number of SGs per cell",
  ) + theme_minimal() + 
  scale_fill_manual(values = manual_colors) + scale_y_continuous(limits = c(0, 1)) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1), legend.position = "bottom", legend.title = element_blank(), axis.title.x = element_blank()) + theme(
         panel.background = element_rect(fill='transparent'),
         plot.background = element_rect(fill='transparent', color=NA),
         panel.grid.major = element_blank(),
         panel.grid.minor = element_blank(),
         legend.background = element_rect(fill='transparent'),
         legend.box.background = element_rect(fill='transparent')
       )

```


### Filter and calculate percentage of cells with SGs
```{r percentage cells with SGs, echo=FALSE}

# Use cells dataset
cells_filtered <- cell[, c("ImageNumber", "ObjectNumber", "Children_SG_G3Bp1_Count", "Children_SG_G3Bp2_Count" , "Children_G3Bp1_G3Bp2_SG_Count")]
cells_filtered <- cells_filtered %>% left_join(Metadata, by = "ImageNumber")

### Create a new data frame for number of cells and number of cells with SGs
new_df <- data.frame(ImageNumber = 1:36,
                     Cell_count = 0,
                     Cells_with_SG_G3Bp1_count = 0,
                     Cells_with_SG_G3Bp2_count = 0,
                     Children_G3Bp1_G3Bp2_SG_Count = 0)

cells_with_SG <- as_tibble(new_df)
rm(new_df)

### Copy number of cells from Image dataset
for(i in 1:dim(cells_with_SG)[1]){
  cells_with_SG$Cell_count[i] <- Image_filtered$Count_cell[i]
}

### Count number of cells with any amount of granules based on G3Bp1
count = 0

for(y in 1:dim(cells_with_SG)[1]){ 
  for(i in 1:dim(cells_filtered)[1]){
    if(cells_filtered$ImageNumber[i] == y){
      if(cells_filtered$Children_SG_G3Bp1_Count[i] > 0){
        count = count + 1
        cells_with_SG$Cells_with_SG_G3Bp1_count[y] <- count}}}
  count <- 0
}


### Count number of cells with any amount of granules based on G3Bp2
count = 0

for(y in 1:dim(cells_with_SG)[1]){ 
  for(i in 1:dim(cells_filtered)[1]){
    if(cells_filtered$ImageNumber[i] == y){
      if(cells_filtered$Children_SG_G3Bp2_Count[i] > 0){
        count = count + 1
        cells_with_SG$Cells_with_SG_G3Bp2_count[y] <- count}}}
  count <- 0
}

### Count number of cells with any amount of granules based on G3Bp1 and G3Bp2
count = 0

for(y in 1:dim(cells_with_SG)[1]){ 
  for(i in 1:dim(cells_filtered)[1]){
    if(cells_filtered$ImageNumber[i] == y){
      if(cells_filtered$Children_G3Bp1_G3Bp2_SG_Count[i] > 0){
        count = count + 1
        cells_with_SG$Children_G3Bp1_G3Bp2_SG_Count[y] <- count}}}
  count <- 0
}

cells_with_SG <- cells_with_SG |> left_join(Metadata, by = "ImageNumber") 
cells_with_SG <- cells_with_SG |> mutate(Ratio_G3Bp1 = Cells_with_SG_G3Bp1_count/Cell_count * 100)
cells_with_SG <- cells_with_SG |> mutate(Ratio_G3Bp2 = Cells_with_SG_G3Bp2_count/Cell_count * 100)
cells_with_SG <- cells_with_SG |> mutate(Ratio_G3Bp1_G3Bp2 = Children_G3Bp1_G3Bp2_SG_Count/Cell_count * 100)
```

### Summarise data per cell_line and Stressor
```{r summarise data per cell_line and Stressor based on percentage of cells with SGs}

cells_with_SG_summary <- cells_with_SG %>%
  group_by(CellLine, Stressor) %>%
  summarise(
    n = n(),                     # Count observations in each group
    mean_Ratio_G3Bp1 = mean(Ratio_G3Bp1),    # Calculate mean Ratio
    sd_Ratio_G3Bp1 = sd(Ratio_G3Bp1),        # Calculate standard deviation
    se_Ratio_G3Bp1 = sd(Ratio_G3Bp1) / sqrt(n()), # Calculate standard error
    mean_Ratio_G3Bp2 = mean(Ratio_G3Bp2),    # Calculate mean Ratio
    sd_Ratio_G3Bp2 = sd(Ratio_G3Bp2),        # Calculate standard deviation
    se_Ratio_G3Bp2 = sd(Ratio_G3Bp2) / sqrt(n()), # Calculate standard error
    mean_Ratio_G3Bp1_G3Bp2 = mean(Ratio_G3Bp1_G3Bp2),    # Calculate mean Ratio
    sd_Ratio_G3Bp1_G3Bp2 = sd(Ratio_G3Bp1_G3Bp2),        # Calculate standard deviation
    se_Ratio_G3Bp1_G3Bp2 = sd(Ratio_G3Bp1_G3Bp2) / sqrt(n()) # Calculate standard error
  ) %>%
  ungroup()
```

### Percentage of cells with stress granules plots
```{r plot data as bars, fig.width=10, fig.height=17}

#change order of labels
cells_with_SG_summary$Stressor <- factor(cells_with_SG_summary$Stressor, 
                                  levels = c("Untreated", "NaAsO2 treated"))

#percentage plot for G3BP1
cells_with_SG_G3Bp1_plot <- cells_with_SG_summary |>
  ggplot(aes(x = CellLine, y = mean_Ratio_G3Bp1, fill = Stressor)) + 
  geom_bar(stat = "identity", position = position_dodge(width = 0.9)) +  
  geom_errorbar(
    aes(ymin = mean_Ratio_G3Bp1 - se_Ratio_G3Bp1, ymax = mean_Ratio_G3Bp1 + se_Ratio_G3Bp1),
    width = 0.2, # Width of the error bar caps
    position = position_dodge(width = 0.9) # Ensure error bars align with dodged bars
  ) +
  labs(
    title = "Cells with G3BP1-positive SGs",
    y = "% cells with SGs",
  ) + theme_minimal() + 
   scale_fill_manual(values = manual_colors) + scale_y_continuous(limits = c(0, 52)) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1), legend.position = "bottom", legend.title = element_blank(), axis.title.x = element_blank()) + theme(
         panel.background = element_rect(fill='transparent'),
         plot.background = element_rect(fill='transparent', color=NA),
         panel.grid.major = element_blank(),
         panel.grid.minor = element_blank(),
         legend.background = element_rect(fill='transparent'),
         legend.box.background = element_rect(fill='transparent')
       )

#percentage plot for G3Bp2
cells_with_SG_G3Bp2_plot <- cells_with_SG_summary |>
  ggplot(aes(x = CellLine, y = mean_Ratio_G3Bp2, fill = Stressor)) + 
  geom_bar(stat = "identity", position = position_dodge(width = 0.9)) +  
  geom_errorbar(
    aes(ymin = mean_Ratio_G3Bp2 - se_Ratio_G3Bp2, ymax = mean_Ratio_G3Bp2 + se_Ratio_G3Bp2),
    width = 0.2, # Width of the error bar caps
    position = position_dodge(width = 0.9) # Ensure error bars align with dodged bars
  ) +
  labs(
    title = "Cells with G3BP2-positive SGs",
    y = "% cells with SG",
  ) + theme_minimal() + 
   scale_fill_manual(values = manual_colors) + scale_y_continuous(limits = c(0, 100)) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1), legend.position = "bottom", legend.title = element_blank(), axis.title.x = element_blank()) + theme(
         panel.background = element_rect(fill='transparent'),
         plot.background = element_rect(fill='transparent', color=NA),
         panel.grid.major = element_blank(),
         panel.grid.minor = element_blank(),
         legend.background = element_rect(fill='transparent'),
         legend.box.background = element_rect(fill='transparent')
       )

#percentage plot for G3BP1/G3Bp2
cells_with_SG_G3Bp12_plot <- cells_with_SG_summary |>
  ggplot(aes(x = CellLine, y = mean_Ratio_G3Bp1_G3Bp2, fill = Stressor)) + 
  geom_bar(stat = "identity", position = position_dodge(width = 0.9)) +  
  geom_errorbar(
    aes(ymin = mean_Ratio_G3Bp1_G3Bp2 - se_Ratio_G3Bp1_G3Bp2, ymax = mean_Ratio_G3Bp1_G3Bp2 + se_Ratio_G3Bp1_G3Bp2),
    width = 0.2, # Width of the error bar caps
    position = position_dodge(width = 0.9) # Ensure error bars align with dodged bars
  ) +
  labs(
    title = "Cells with G3Bp1/G3Bp2-positive SGs",
    y = "% cells with SG",
  ) + theme_minimal() + 
   scale_fill_manual(values = manual_colors) + scale_y_continuous(limits = c(0, 50)) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1), legend.position = "bottom", legend.title = element_blank(), axis.title.x = element_blank()) + theme(
         panel.background = element_rect(fill='transparent'),
         plot.background = element_rect(fill='transparent', color=NA),
         panel.grid.major = element_blank(),
         panel.grid.minor = element_blank(),
         legend.background = element_rect(fill='transparent'),
         legend.box.background = element_rect(fill='transparent')
       )
```